### 2015-12-08 paper/4-performance.tex
Updated citation to HighInterconnect and fixed some float options to make it compile on a minimal latex install

\subsubsection{Interconnect}
Interconnect~\cite{van2013research} is a benchmark where all models broadcast, creating a complete graph in dependencies between models. As the model count increases, we see (see Figure~\ref{fig:Interconnect} the expected quadratic increase in runtime for both adevs and dxex, but an increasing penalty for dxex. Profiling shows this is entirely due to the heap allocation of messages, which even though minimized by using memory pools remains significant.
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig3.eps}
	\label{fig3.eps}
	\caption{Interconnect}
	\label{fig:Interconnect}
\end{figure}

\subsection{Parallel Simulation}
%TODO explain that we will now compare different synchronization protocols
%TODO compare different synchronization protocols with adevs' conservative, showing that we are faster (or at least, not that much slower) than adevs
By default, the benchmarks use 4 kernels for parallel simulation.
\subsubsection{Devstone}
The flattened models are allocated to kernels by giving each kernel a distinct section of the chain, resulting in a low ratio of inter-kernel to intra-kernel messages. For the optimistic case, this can cause more reverts since the kernels will start to drift faster as the model-count increases. Furthermore, optimistic synchronization is quite sensitive to an increase in kernels, since the delay before a revert propagates, increases. This benchmark requires a specific warm-up time; for $n=d\times w$ models, it takes $\textup{timeadvance}()*(n-1)$ transitions to activate the last model in the chain. When using parallel processing, this can be reduced to $\textup{timeadvance}()*(n\frac{kernels-1}{kernels} -1)$ before the last kernel becomes active.\\
Figure~\ref{fig:DevstoneParallel} shows that dxex outperformance adevs (in both optimistic and conservative case).
\begin{figure}[ltbh]
	\includegraphics[width=.5\textwidth]{fig/fig2.eps}
	\label{fig:DevstoneParallel}
	\caption{DevStone parallel}
\end{figure}
\subsubsection{PHold}
%Phold
In Phold, the allocation is specified in the benchmark itself. Each kernel manages a single node with a constant set of sub-nodes. The parameter $R$ (which is here set to $10$) determines the percentage of remote destination models.\\
The dynamic dependency graph is a very sparse version of the static dependency graph, penalizing the conservative case. The lookahead is $\epsilon$, so the conservative case spends most of its time crawling in steps of $\epsilon$. Since the dependency graph between kernels is a complete graph, this is not a simulation that scales in our implementation. For $N$ kernels, each kernel has to query the null-time of $N-1$ kernels, resulting in O($N^2$) polling behaviour. In a non-cyclic simulation with a non-trivial lookahead (like in, e.g., Devstone), that choice does pay off (see Figure~\ref{fig:DevstoneParallel}). Adevs's lesser performance is due in part to their lookahead management, which after profiling shows to spend a non-trivial amount of time in exception handling code.\\
The optimistic case suffers little from the above problems; due to the high interconnectivity, however, a cascading revert is still possible. With the percentage of remotes equal to $100$, Phold reflects interconnect in behaviour, which is why the $R$ value is not dimensioned here. A revert is very expensive in PHold due to our usage of C++11's random nr generators. The cost of a revert is dominated by the recalculation of destination models, not in allocating/deallocating states/messages. Once a revert happens the drift between kernels increases fast, increasing the likelihood of more reverts. Despite all this, optimistic can for low $R$ values quickly exploit the uncertainty that slows down conservative in this benchmark.

@@

\subsubsection{Priority network model}
%TODO show a single plot which varies a parameter that changes the ideal synchronization protocol
The priority benchmark is composed of a single server generating a stream of $0\leq m \leq n$ messages at fixed time intervals, interleaved with a probability $p$ for a priority message to $n$ receivers (see Figure~\ref{fig:NetworkModel}).\\ This defaults the lookahead for the receivers to $\epsilon$, but this time there is no scaling effect, nor are there cycles in the dependency graph. This model therefore highlights the basic strengths/weaknesses of both synchronization protocols. Receiving models are allocated on another kernel than the server and have an internal transition so that they will not wait for the incoming messages.
\begin{center}
\begin{figure}
\input pqueue.tex
\caption{The priority network model}
\label{fig:NetworkModel}
\end{figure}
\end{center}
A key difference here with the other benchmarks is that a state (in the Receiver instances) is very cheap to copy/create. The kernel holding the server will never revert since it is a source in the dependency graph. The optimistic case will therefore not suffer the same performance hit in recreating the states as it does in PHold. The overhead in the optimistic case is entirely due to the factor $m$, which will quickly dominate in increasing buffers of received messages.
\begin{figure}[ltbh]
	\includegraphics[width=.5\textwidth]{fig/fig5.eps}
	\label{fig:Priority}
	\caption{Priority}
\end{figure}
Interestingly, the parameter $p$ does not clearly favour either synchronization protocol (see Figure~\ref{fig:Priority}). While this removes any possibility for a lookahead, the conservative case can quickly bridge the timespan between fixed messages since there is no cyclic dependency. Lightweight states prevent performance loss due to reverts in the optimistic case; only the overhead in event handling in the optimistic case eventually becomes the deciding factor.

@@

\begin{table}[lhtb]

@@

\paragraph*{Priority Network}
The priority network model is detailed in section 4.1.\\
The high memory usage of the conservative case is largely due to the garbage collection implementation. Kernel 0 is the initiator of the LBTS calculation but since it holds only models independent of any other (server), it will finish simulation very fast leaving the other kernel bereft of an updated LBTS. The kernel holding the server will after simulation wait on the receiving kernel until it can prove all sent messages can be deallocated.
\begin{table}[lhtb]
	\centering
	\label{pmod_mem}
	\begin{tabular}{| l | l | l |}
		\hline
		dxex &dxex con &dxex opt\\ \hline
		35 & 450 & 651\\ \hline
	\end{tabular}
	\caption{Priority model n 128, m 16, p 10,  t2e8, unit MiB, 2 kernels}
\end{table}

>>>>

\subsubsection{Interconnect}
Interconnect~\cite{MasterThesis} is a benchmark where all models broadcast, creating a complete graph in dependencies between models. As the model count increases, we see (see Figure~\ref{fig:Interconnect} the expected quadratic increase in runtime for both adevs and dxex, but an increasing penalty for dxex. Profiling shows this is entirely due to the heap allocation of messages, which even though minimized by using memory pools remains significant.
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig3.eps}
	\label{fig3.eps}
	\caption{Interconnect}
	\label{fig:Interconnect}
\end{figure}

\subsection{Parallel Simulation}
%TODO explain that we will now compare different synchronization protocols
%TODO compare different synchronization protocols with adevs' conservative, showing that we are faster (or at least, not that much slower) than adevs
By default, the benchmarks use 4 kernels for parallel simulation.
\subsubsection{Devstone}
The flattened models are allocated to kernels by giving each kernel a distinct section of the chain, resulting in a low ratio of inter-kernel to intra-kernel messages. For the optimistic case, this can cause more reverts since the kernels will start to drift faster as the model-count increases. Furthermore, optimistic synchronization is quite sensitive to an increase in kernels, since the delay before a revert propagates, increases. This benchmark requires a specific warm-up time; for $n=d\times w$ models, it takes $\textup{timeadvance}()*(n-1)$ transitions to activate the last model in the chain. When using parallel processing, this can be reduced to $\textup{timeadvance}()*(n\frac{kernels-1}{kernels} -1)$ before the last kernel becomes active.\\
Figure~\ref{fig:DevstoneParallel} shows that dxex outperformance adevs (in both optimistic and conservative case).
\begin{figure}
	\includegraphics[width=.5\textwidth]{fig/fig2.eps}
	\label{fig:DevstoneParallel}
	\caption{DevStone parallel}
\end{figure}
\subsubsection{PHold}
%Phold
In Phold, the allocation is specified in the benchmark itself. Each kernel manages a single node with a constant set of sub-nodes. The parameter $R$ (which is here set to $10$) determines the percentage of remote destination models.\\
The dynamic dependency graph is a very sparse version of the static dependency graph, penalizing the conservative case. The lookahead is $\epsilon$, so the conservative case spends most of its time crawling in steps of $\epsilon$. Since the dependency graph between kernels is a complete graph, this is not a simulation that scales in our implementation. For $N$ kernels, each kernel has to query the null-time of $N-1$ kernels, resulting in O($N^2$) polling behaviour. In a non-cyclic simulation with a non-trivial lookahead (like in, e.g., Devstone), that choice does pay off (see Figure~\ref{fig:DevstoneParallel}). Adevs's lesser performance is due in part to their lookahead management, which after profiling shows to spend a non-trivial amount of time in exception handling code.\\
The optimistic case suffers little from the above problems; due to the high interconnectivity, however, a cascading revert is still possible. With the percentage of remotes equal to $100$, Phold reflects interconnect in behaviour, which is why the $R$ value is not dimensioned here. A revert is very expensive in PHold due to our usage of C++11's random nr generators. The cost of a revert is dominated by the recalculation of destination models, not in allocating/deallocating states/messages. Once a revert happens the drift between kernels increases fast, increasing the likelihood of more reverts. Despite all this, optimistic can for low $R$ values quickly exploit the uncertainty that slows down conservative in this benchmark.

@@

\subsubsection{Priority network model}
%TODO show a single plot which varies a parameter that changes the ideal synchronization protocol
The priority benchmark is composed of a single server generating a stream of $0\leq m \leq n$ messages at fixed time intervals, interleaved with a probability $p$ for a priority message to $n$ receivers (see Figure~\ref{fig:NetworkModel}).\\ This defaults the lookahead for the receivers to $\epsilon$, but this time there is no scaling effect, nor are there cycles in the dependency graph. This model therefore highlights the basic strengths/weaknesses of both synchronization protocols. Receiving models are allocated on another kernel than the server and have an internal transition so that they will not wait for the incoming messages.
\begin{center}
\begin{figure}
\input pqueue.tex
\caption{The priority network model}
\label{fig:NetworkModel}
\end{figure}
\end{center}
A key difference here with the other benchmarks is that a state (in the Receiver instances) is very cheap to copy/create. The kernel holding the server will never revert since it is a source in the dependency graph. The optimistic case will therefore not suffer the same performance hit in recreating the states as it does in PHold. The overhead in the optimistic case is entirely due to the factor $m$, which will quickly dominate in increasing buffers of received messages.
\begin{figure}
	\includegraphics[width=.5\textwidth]{fig/fig5.eps}
	\label{fig:Priority}
	\caption{Priority}
\end{figure}
Interestingly, the parameter $p$ does not clearly favour either synchronization protocol (see Figure~\ref{fig:Priority}). While this removes any possibility for a lookahead, the conservative case can quickly bridge the timespan between fixed messages since there is no cyclic dependency. Lightweight states prevent performance loss due to reverts in the optimistic case; only the overhead in event handling in the optimistic case eventually becomes the deciding factor.

@@

\begin{table}

@@

\paragraph*{Priority Network}
The priority network model is detailed in section 4.1.\\
The high memory usage of the conservative case is largely due to the garbage collection implementation. Kernel 0 is the initiator of the LBTS calculation but since it holds only models independent of any other (server), it will finish simulation very fast leaving the other kernel bereft of an updated LBTS. The kernel holding the server will after simulation wait on the receiving kernel until it can prove all sent messages can be deallocated.
\begin{table}
	\centering
	\label{pmod_mem}
	\begin{tabular}{| l | l | l |}
		\hline
		dxex &dxex con &dxex opt\\ \hline
		35 & 450 & 651\\ \hline
	\end{tabular}
	\caption{Priority model n 128, m 16, p 10,  t2e8, unit MiB, 2 kernels}
\end{table}
